Mastery Learning in Practice:\\ A (Mostly) Descriptive Analysis Based on the Cognitive Tutor Algebra I Effectiveness Trial

### 2017-12-05 paper.Rnw
couple tweaks
added a "total" line to one of the tables and fixed the formatting on
table 1

There are some limitations to the available data for this study.
Log data from some schools, and some students within schools, were
missing either because the log files were not retrievable, or because
of an imperfect ability to link log data to other study data files.
For this reason, this study uses only data from on the
\Sexpr{n_distinct(data[['schoolid2']])} schools for which at least 80\%
students in both study years appear in the log data file.
This sample includes \Sexpr{n_distinct(data[['field_id']])} students, around
\Sexpr{round(n_distinct(data[['field_id']])/n_distinct(stud[['field_id']]))*100}\%
of the total high-school sample.
Table \ref{tab:nByState} gives the number of students in each state of
the sample in each year.
The states in the table are ordered by the total number of students
they represent in the sample; they will appear in this order in
all of the forthcoming tables and figures.
Some figures will only show data from a subset of states; since so few
students were in New Jersey, it will be excluded from almost all
state-by-state comparisons (but included analyses that pool across states).
<<nByState,dependson='data'>>=
xtable(data%>%group_by(state,year)%>%summarize(n=n_distinct(field_id))%>%dcast(year~state),
       caption='Numbers of students in the sample by state and study year',label='tab:nByState')
@

@@

tab1$variable <- as.character(tab1$variable)
tab2$variable <- as.character(tab2$variable)
cat('&')
cat(names(tab1)[-1],sep='&')
cat('\\\\')
for(i in 1:nrow(tab1)){
    cat(tab1[i,1],round(unlist(tab1[i,-1])),sep='&')
    cat('\\\\')
}
cat('&\\multicolumn{6}{c}{Year 2}\\\\ \n')
for(i in 1:nrow(tab2)){
    cat(tab2[i,1],round(unlist(tab2[i,-1])),sep='&')
    cat('\\\\')
}
rownames(tab1) <- rownames(tab2) <- tab1$variable
tab1 <- as.matrix(tab1[,-1])
tab2 <- as.matrix(tab2[,-1])

>>>>

There are some limitations to the available data for this study.
Log data from some schools, and some students within schools, were
missing either because the log files were not retrievable, or because
of an imperfect ability to link log data to other study data files.
For this reason, this study uses only data from on the
\Sexpr{n_distinct(data[['schoolid2']])} schools for which at least 80\%
students in both study years appear in the log data file.
This sample includes \Sexpr{n_distinct(data[['field_id']])} students, around
\Sexpr{round(n_distinct(data[['field_id']])/n_distinct(stud[['field_id']]))*100}\%
of the total high-school sample.
Table \ref{tab:nByState} gives the number of students in each state of
the sample in each year.
The states in the table are ordered by the total number of students
they represent in the sample; they will appear in this order in
all of the forthcoming tables and figures.
Some figures will only show data from a subset of states; since so few
students were in New Jersey, it will be excluded from almost all
state-by-state comparisons (but included analyses that pool across states).
<<nByState,dependson='data'>>=
sampleSizeTable <- data%>%group_by(state,year)%>%summarize(n=n_distinct(field_id))%>%dcast(year~state)
sampleSizeTable <- sampleSizeTable[,-1]
rownames(sampleSizeTable) <- c('Year 1','Year 2')

print(xtable(sampleSizeTable,caption='Numbers of students in the sample by state and study year',label='tab:nByState'),include.rownames=TRUE)
@

@@

tab1$variable <- as.character(tab1$variable)
tab2$variable <- as.character(tab2$variable)
cat('&')
cat(names(tab1)[-1],sep='&')
cat('\\\\')
for(i in 1:nrow(tab1)){
    cat(tab1[i,1],round(unlist(tab1[i,-1])),sep='&')
    cat('\\\\')
}
cat('Total',round(colSums(tab1[,-1])),sep='&')
cat('\\\\')
cat('&\\multicolumn{6}{c}{Year 2}\\\\ \n')
for(i in 1:nrow(tab2)){
    cat(tab2[i,1],round(unlist(tab2[i,-1])),sep='&')
    cat('\\\\')
}
cat('Total',round(colSums(tab2[,-1])),sep='&')
cat('\\\\')

rownames(tab1) <- rownames(tab2) <- tab1$variable
tab1 <- as.matrix(tab1[,-1])
tab2 <- as.matrix(tab2[,-1])

### 2017-12-05 paper.tex
couple tweaks
added a "total" line to one of the tables and fixed the formatting on
table 1

There are some limitations to the available data for this study.
Log data from some schools, and some students within schools, were
missing either because the log files were not retrievable, or because
of an imperfect ability to link log data to other study data files.
For this reason, this study uses only data from on the
18 schools for which at least 80\%
students in both study years appear in the log data file.
This sample includes 4460 students, around
100\%
of the total high-school sample.
Table \ref{tab:nByState} gives the number of students in each state of
the sample in each year.
The states in the table are ordered by the total number of students
they represent in the sample; they will appear in this order in
all of the forthcoming tables and figures.
Some figures will only show data from a subset of states; since so few
students were in New Jersey, it will be excluded from almost all
state-by-state comparisons (but included analyses that pool across states).
% latex table generated in R 3.3.1 by xtable 1.8-2 package
% Mon Nov  6 13:55:49 2017
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & year & TX & KY & MI & LA & CT & NJ \\ 
  \hline
1 &   1 & 947 & 890 & 325 & 164 & 112 &  17 \\ 
  2 &   2 & 719 & 646 & 285 & 180 & 139 &  36 \\ 
   \hline
\end{tabular}
\caption{Numbers of students in the sample by state and study year} 
\label{tab:nByState}
\end{table}

@@

% latex table generated in R 3.3.1 by xtable 1.8-2 package
% Mon Nov  6 13:58:09 2017
\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Hours & Problems & Sections & Units \\ 
  \hline
Year 1 & 33.47 & 309 & 43.0 & 9 \\ 
  Year 2 & 23.83 & 228 & 36.5 & 10 \\ 
   \hline
\end{tabular}
\caption{Median numbers of hours, problems, sections, and units worked by each student in the dataset in the two years of the study. Students with no usage data were excluded.} 
\label{tab:medUsage}
\end{table}

@@

&   \multicolumn{6}{c}{Year 1}\\%&\multicolumn{6}{c}{Year 2}\\
&TX&KY&MI&LA&CT&Overall\\Final&699&625&242&129&91&1796\\Reassigned&600&140&220&5&92&1082\\Promoted&1619&9628&4582&1459&261&17561\\Mastered&15815&56199&31200&8648&1303&113374\\&\multicolumn{6}{c}{Year 2}\\ 
Final&1252&660&236&137&114&2412\\Reassigned&1215&253&275&9&144&2042\\Promoted&1000&6296&4717&1998&450&14477\\Mastered&4929&36676&26509&10108&1513&79874\\

@@

\begin{table}
  \centering
  \begin{tabular}{rll|ll}
    &\multicolumn{2}{c}{Year 1}&\multicolumn{2}{c}{Year 2}\\
    &Aug--Dec&Jan--Jun&Aug--Dec&Jan--Jun\\
% latex table generated in R 3.3.1 by xtable 1.8-2 package
% Mon Nov  6 14:19:52 2017
 TX &  18 & 582 & 281 & 932 \\ 
  KY &   9 & 131 &  52 & 201 \\ 
  MI &  47 & 173 & 245 &  30 \\ 
  LA &   0 &   5 &   9 &   0 \\ 
  CT &  14 &  78 &  84 &  60 \\ 
  NJ &   3 &  22 &  17 & 129 \\ 
   \hline
Overall &  91 & 991 & 688 & 1352 \\

>>>>

There are some limitations to the available data for this study.
Log data from some schools, and some students within schools, were
missing either because the log files were not retrievable, or because
of an imperfect ability to link log data to other study data files.
For this reason, this study uses only data from on the
18 schools for which at least 80\%
students in both study years appear in the log data file.
This sample includes 4460 students, around
100\%
of the total high-school sample.
Table \ref{tab:nByState} gives the number of students in each state of
the sample in each year.
The states in the table are ordered by the total number of students
they represent in the sample; they will appear in this order in
all of the forthcoming tables and figures.
Some figures will only show data from a subset of states; since so few
students were in New Jersey, it will be excluded from almost all
state-by-state comparisons (but included analyses that pool across states).
% latex table generated in R 3.4.2 by xtable 1.8-2 package
% Tue Dec 05 09:39:47 2017
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrr}
  \hline
 & TX & KY & MI & LA & CT & NJ \\ 
  \hline
Year 1 & 947 & 890 & 325 & 164 & 112 &  17 \\ 
  Year 2 & 719 & 646 & 285 & 180 & 139 &  36 \\ 
   \hline
\end{tabular}
\caption{Numbers of students in the sample by state and study year} 
\label{tab:nByState}
\end{table}

@@

% latex table generated in R 3.4.2 by xtable 1.8-2 package
% Tue Dec 05 09:40:25 2017
\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Hours & Problems & Sections & Units \\ 
  \hline
Year 1 & 33.47 & 309 & 43.0 & 9 \\ 
  Year 2 & 23.83 & 228 & 36.5 & 10 \\ 
   \hline
\end{tabular}
\caption{Median numbers of hours, problems, sections, and units worked by each student in the dataset in the two years of the study. Students with no usage data were excluded.} 
\label{tab:medUsage}
\end{table}

@@

&   \multicolumn{6}{c}{Year 1}\\%&\multicolumn{6}{c}{Year 2}\\
&TX&KY&MI&LA&CT&Overall\\Final&699&625&242&129&91&1796\\Reassigned&600&140&220&5&92&1082\\Promoted&1619&9628&4582&1459&261&17561\\Mastered&15815&56199&31200&8648&1303&113374\\Total&18733&66592&36244&10241&1747&133813\\&\multicolumn{6}{c}{Year 2}\\ 
Final&1252&660&236&137&114&2412\\Reassigned&1215&253&275&9&144&2042\\Promoted&1000&6296&4717&1998&450&14477\\Mastered&4929&36676&26509&10108&1513&79874\\Total&8396&43885&31737&12252&2221&98805\\

@@

\begin{table}
  \centering
  \begin{tabular}{rll|ll}
    &\multicolumn{2}{c}{Year 1}&\multicolumn{2}{c}{Year 2}\\
    &Aug--Dec&Jan--Jun&Aug--Dec&Jan--Jun\\
% latex table generated in R 3.4.2 by xtable 1.8-2 package
% Tue Dec 05 10:03:13 2017
 TX &  18 & 582 & 281 & 932 \\ 
  KY &   9 & 131 &  52 & 201 \\ 
  MI &  47 & 173 & 245 &  30 \\ 
  LA &   0 &   5 &   9 &   0 \\ 
  CT &  14 &  78 &  84 &  60 \\ 
  NJ &   3 &  22 &  17 & 129 \\ 
   \hline
Overall &  91 & 991 & 688 & 1352 \\
